\chapter{Theoretical Basis}

\section{Background Theory}%
\label{sec:background_theory}

\subsubsection{Clock and Data Recovery}%
\label{ssub:clock_and_data_recovery}
Commonly a serial data stream is sent over a channel without a clock signal.
Clock recovery is the process of extracting timing information from a serial
data stream. This timing data is then used to decode the received data stream.
This process is known as Clock and Data Recovery (CDR).

In this application a bang-bang CDR circuit is used. 
In this case transitions in a received data signal are counted as "early" or
"late" as compared with a local clock. The clock can then be adjusted based on
the local transitions~\cite{alexander1975clock}.
\TODO{Image of bang-bang CDR}

\subsubsection{Source Synchronous System}%
\label{ssub:source_synchronous_system}
In a source synchronous system a clock signal is provided alongside the data
signal. 

\subsubsection{PRBS Generation}%
\label{ssub:prbs_generation}
A pseudorandom binary sequence (PRBS) is a sequence of bits that appears to be
random. However as it is generated using a deterministic algorithim, it can be
replicated if the inital conditions are the same.

\section{Literature Review}%
\label{sec:literature_review}

\noindent \cite{kari_phase} outlines how CDR circuits are a limiting factor in
optical switching and proposes a method of overcoming this (phase caching).
Through phase caching they were able to demonstrate sub nanosecond locking
times, improving the locking time 12x. This is a different approach that seeks
to solve the same problem.

\noindent In \cite{serrano2013white}, \cite{moreira2010digital}, and
\cite{moreira2009white}, the white rabbit project is discussed. A white rabbit
system provides sub-nanosecond synchronisation accuracy. To achieve this,
accurate measurements of the link delay between the nodes of the network must
be calculated.  While instructive, the method is not directly applicable to
the project, as in a White Rabbit system, all the nodes are locked to the same
frequency. Hence the link delay can be calculated by having a node receive a
clock signal from another node, then return the same signal. The link delay can
then be calculated by comparing the phase offset of the two signals.

\noindent \cite{williams2016source} described an optical source synchronous
system. It describes how choosing the correct wavelength for the clock can
minimise the modal cross-talk. Furthermore, in conjunction with
\cite{ragab2011receiver} it describes how source synchronous systems are able
to track correlated jitter between clock and data channels, and how system
performance can be degraded by channel slew between clock and data channels.\\
\cite{williams2019reconfiguration} further explored reducing the modal
crosstalk by proposing an architecture with re-configurable clock and data
paths, thus allowing the user to chose the optimal lane for the sensitive clock
for each photonic interconnect. This may not be needed however, as each
transmitter should have a fixed data characteristic.

\noindent \cite{chen2017optimization} and \cite{fixed_latency} describe fixed
latency links. In the event we were unable to bypass the CDR, it may be
possible to organise the system to have a fixed latency, then force the CDR to
the appropriate fixed phase. Thus the circuit could thus have a much reduced
CDR lock time.

\noindent \cite{dru_guide}, \cite{nidru} describe an intellectual property (IP)
that allows the high speed serial transceivers to be used at much lower data
rates. This was initially of interest because it would have been easier to
demonstrate a working system with lower data rates. However as this is an extra
IP used in conjunction with the transceivers it did not turn out to be useful
for the project. 

\noindent \cite{mendes_transceiver} this presentation describes a system where
the phase of a transceiver on Xilinx board is kept stable over resets. While
this was done primarily on the transmitter side this still shows that fixing
the phase of the transceiver is possible.

